Congestion-Free Network Updates: Algorithms and Complexity

### 2023-05-01 paper/main.tex
Add section on permutations (of blocks)

% Created 2023-04-16 Sun 15:13
% Intended LaTeX compiler: pdflatex
\documentclass[fontsize=11pt,paper=a4]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{preamble}
\author{Mitja Daniel Krebs}
\date{\today}
\title{Congestion-Free Network Updates: Algorithms and Complexity}
\hypersetup{
 pdfauthor={Mitja Daniel Krebs},
 pdftitle={Congestion-Free Network Updates: Algorithms and Complexity},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 28.2 (Org mode 9.6.1)}, 
 pdflang={English}}
\begin{document}

@@

\begin{table}[htbp]
\caption{Table of notations}
\centering
\begin{tabular}{ll}
\hyperref[org1b69eba]{\(\leq_G\)} & The reachability relation for directed graph \(G\)\\[0pt]
\hyperref[org072524f]{\(b(v,P)\)} & The \(P\)-block containing vertex \(v\)\\[0pt]
\hyperref[org51c0575]{\(P(b)\)} & The flow pair \(P\) such that block \(b\) is a \(P\)-block\\[0pt]
\hyperref[orgdcabd12]{\(\mathcal{S}(b)\)} & The least vertex in block \(b\) w.r.t. reachability relation \(\leq_{P(b)^o\cup P(b)^u}\)\\[0pt]
\hyperref[org3a72523]{\(B^P(G)\)} & The set of \(P\)-blocks\\[0pt]
\hyperref[orgaf5c11c]{\(B(G)\)} & The set of blocks\\[0pt]
\hyperref[org1d57970]{\(\mathcal{B}(b)\)} & The round in which block \(b\) is updated\\[0pt]
\hyperref[org45f406f]{\(\mathcal{B}(v,P)\)} & The round in which block \(b(v,P)\) is updated\\[0pt]
\hyperref[orge41a594]{\(B_i\)} & The set of blocks updated before or in the \(i\)-th round\\[0pt]
 & \\[0pt]
\end{tabular}
\end{table}

\part{Preliminaries}
\label{sec:org08c7ac8}

\begin{notation}
For a directed graph \(G\), \(\leq_G=E(G)^*\) denotes the reachability relation of \(G\).
That is, for every two vertices \(u,v\in V(G)\), \(u\leq_Gv\) iff there is a path in \(G\) from \(u\) to \(v\).
\label{org1b69eba}
\end{notation}

@@

\begin{defn}
Let \(G=(V,E,\mathcal{P},s,t,c)\) be an update flow network and \(P\in\mathcal{P}\) be a flow pair.
Let \(v_1,\dots,v_{\ell}\) be the set \(V(P^o\cap P^u)\) ordered w.r.t. \(\leq_{P^o\cup P^u}\).
For every \(i\in[\ell-1]\), we define the \(i\)-th \emph{\(P\)-block} as \(b_i^P=\{v\mid v_i\leq_{P^o\cup P^u}v\leq_{P^o\cup P^u}v_{i+1}\}\).
\label{org3e1b5bf}
\end{defn}

@@

\begin{notation}
For a flow pair \(P\) and a vertex \(v\in V(P)\), \(b(v,P)\) denotes the \(P\)-block containing \(v\).
\label{org072524f}
\end{notation}

\begin{notation}
For a block \(b\), \(P(b)\) denotes the flow pair \(P\) such that \(b\) is a \(P\)-block.
\label{orgdcabd12}
\end{notation}

\begin{notation}
For a block \(b\), \(\mathcal{S}(b)\) denotes the \emph{start} of \(b\), that is, the least vertex in \(b\) w.r.t. \(\leq_{P(b)^o\cup P(b)^u}\).
\label{org51c0575}
\end{notation}

\begin{notation}
For an update flow network \(G\) and a flow pair \(P\), \(B^P(G)\) denotes the set of \(P\)-blocks.
\label{org3a72523}
\end{notation}

\begin{notation}
For an update flow network \(G\), \(B(G)=\bigcup_{P\in\mathcal{P}}B^P(G)\) denotes the set of blocks.
\label{orgaf5c11c}
\end{notation}

\begin{defn}
A \emph{block sequence} \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) is an ordered partition of the set of blocks.
\label{org237f83d}
\end{defn}

\begin{remark}
We may ignore all blocks containing less than three vertices.
\label{org7d3303c}
\end{remark}

@@

\begin{notation}
For a block sequence \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) and a block \(b\), \(\mathcal{B}(b)\) denotes the index \(i\in[\ell]\) such that \(b\) is contained in \(\mathscr{B}_i\).
\label{org1d57970}
\end{notation}

\begin{notation}
For a block sequence \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\), a flow pair \(P\), and a vertex \(v\in V(P)\), \(\mathcal{B}(v,P)=\mathcal{B}(b(v,P))\) denotes the index \(i\in[\ell]\) such that block \(b(v,P)\) is contained in \(\mathscr{B}_i\).
\label{org45f406f}
\end{notation}

\begin{notation}
For a block sequence \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) and an index \(i\in[\ell]\), \(B_i=\bigcup_{j\leq i}\mathscr{B}_j\) denotes the set of blocks updated before or in the \(i\)-th round.
\label{orge41a594}
\end{notation}

\begin{defn}
Let \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) be a block sequence.
For a flow pair \(P\), an edge \((u,v)\in E(P^o\cup P^u)\), and an index \(i\in[\ell]\), the \emph{activation label} \(\alpha_P((u,v),B_i)\) is defined as follows:
\[\alpha_P((u,v),B_i)=
\begin{cases}
\mathrm{active} & \text{if }(u,v)\in E(P^o)\text{ and }b(u,P)\notin B_i\\
\mathrm{active} & \text{if }(u,v)\in E(P^u)\text{ and }b(u,P)\in B_i\\
\mathrm{inactive} & \text{otherwise}.
\end{cases}\]
\label{org4ed0822}
\end{defn}

@@

\item If \((u,v)\in E(P^u\setminus P^o)\), then
\[\alpha_P((u,v),B_i)=
   \begin{cases}
   \mathrm{active} & i\geq\mathcal{B}(u,P)\\
   \mathrm{inactive} & i<\mathcal{B}(u,P).
   \end{cases}\]
\end{enumerate}
\label{orgcb6370f}
\end{lem}

\begin{notation}
For a flow pair \(P\) and a \(P\)-block \(b\), \(U(b)=\{(v,P)\mid v\in b\}\) denotes the set of updates induced by \(b\).
Moreover, for a set \(B\) of blocks, \(U(B)=\bigcup_{b\in B}U(b)\) denotes the set of updates induced by \(B\).
\label{orgc651e82}
\end{notation}

@@

\begin{lem}
Let \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) be a block sequence, \(P\) be a flow pair, \(e\in E(P^o\cup P^u)\), and \(i\in[\ell]\).
Then \(\alpha_P(e,B_i)=\mathrm{active}\) iff \(e\in E(T_{P,U(B_i)})\).
\label{orgc473ad3}
\end{lem}

\begin{defn}
A block sequence \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) is \emph{feasible} if for every edge \(e\) and every index \(i\in[\ell]\),
\begin{equation}
\label{eq:org8214b0f}
c(e)\geq\sum_{P\in\mathcal{P}:\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}}d_P,
\end{equation}
where we define \(\mathscr{B}_0\) to be the empty set.
\label{org8171bc3}
\end{defn}

\begin{remark}
Let \(G\) be an update flow network with unit demand, that is, \(d_P=1\) for every flow pair \(P\), and let \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) be a block sequence.
Then, for every edge \(e\) and every index \(i\in[\ell]\), capacity constraint \ref{eq:org8214b0f} simplifies to:
\begin{align*}
c(e)
&\geq\sum_{P\in\mathcal{P}:\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}}d_P\\
&=\sum_{P\in\mathcal{P}:\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}}1\\
&=\lvert\{P\in\mathcal{P}\mid\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}\}\rvert.
\end{align*}
\label{org4411b52}
\end{remark}

@@

\begin{enumerate}
\item \label{itm:lem-update-flow-network-feasible-if-1}
The old flow network is feasible if capacity constraint \ref{eq:org8214b0f} is satisfied for every edge and \(i=1\).

\item \label{itm:lem-update-flow-network-feasible-if-2}
The updated flow network is feasible if capacity constraint \ref{eq:org8214b0f} is satisfied for every edge and \(i=\ell\).
\end{enumerate}
\label{org1280734}
\end{lem}

@@

\paragraph{\ref{itm:lem-update-flow-network-feasible-if-1}.}
Suppose capacity constraint \ref{eq:org8214b0f}  is satisfied for \(e\) and \(i=1\).
Then, since \(\mathscr{B}_0=\emptyset\), and by definitions of \(B_i\) and \(\alpha_P\):
\begin{align*}
c(e)
&\geq\sum_{P\in\mathcal{P}:\alpha_P(e,B_0)=\mathrm{active}\text{ or }\alpha_P(e,B_1)=\mathrm{active}}d_P\\
&\geq\sum_{P\in\mathcal{P}:\alpha_P(e,B_0)=\mathrm{active}}d_P\\
&=\sum_{P\in\mathcal{P}:e\in E(P^o)}d_P.
\end{align*}
\paragraph{\ref{itm:lem-update-flow-network-feasible-if-2}.}
Suppose capacity constraint \ref{eq:org8214b0f}  is satisfied for \(e\) and \(i=\ell\).
Then, since \(\mathcal{B}\) partitions the set of blocks, and by definitions of \(B_i\) and \(\alpha_P\):
\begin{align*}
c(e)
&\geq\sum_{P\in\mathcal{P}:\alpha_P(e,B_{\ell-1})=\mathrm{active}\text{ or }\alpha_P(e,B_{\ell})=\mathrm{active}}d_P\\
&\geq\sum_{P\in\mathcal{P}:\alpha_P(e,B_{\ell})=\mathrm{active}}d_P\\
&=\sum_{P\in\mathcal{P}:e\in E(P^u)}d_P.
\end{align*}
\end{proof}

\begin{corollary}
There is a feasible block sequence iff there is a feasible update sequence.
\label{org8c7cbdd}
\end{corollary}

\part{\(\textbf{NP}\)-Hardness for \(k=3\)}
\label{sec:org21620c8}

@@

\begin{thm}
The \(k\)-network flow update problem is \(\textbf{NP}\)-hard for \(k=3\).
\label{org2706698}
\end{thm}

@@

\begin{thm}
The \(k\)-network flow update problem, where every edge is used by at most three flow pairs, is \(\textbf{NP}\)-hard for \(k=10\).
\label{org65ec49d}
\end{thm}

Then, we will (repeatedly) apply the following lemma to the flow update network we will have constructed in the proof of Theorem \ref{org65ec49d} to reduce the number of flow pairs from \(10\) to \(3\).

@@

Then there is an update flow network \(\tilde{G}\) with \(k-1\) flow pairs such that (\(\lvert\tilde{G}\rvert=O(\lvert G\rvert)\) and) there is a feasible block sequence \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) for \(G\) iff there is a feasible block sequence \(\tilde{\mathcal{B}}=(\tilde{\mathscr{B}}_1,\dots,\tilde{\mathscr{B}}_{\ell})\) for \(\tilde{G}\).
\label{orgce7cb77}
\end{lem}

@@

\chapter{\(\textbf{NP}\)-Hardness for the Special Case}
\label{sec:org4acbd6a}

The proof of Theorem \ref{org65ec49d} is via reduction from \(\textsf{4-SAT}\) and is based on the \(\textbf{NP}\)-hardness proof for \(k=6\) in (Amiri, Saeed A. and Dudycz, Szymon and Parham, Mahmoud and Schmid, Stefan and Wiederrecht, Sebastian, 2019).

@@

\section{The Reduction}
\label{sec:org9ac37d7}

@@

We now proceed with the detailed specification of clause gadget \(C^i\) (see Figure \ref{fig:org4c8ec8f}).

>>>>

% Created 2023-05-01 Mon 11:38
% Intended LaTeX compiler: pdflatex
\documentclass[fontsize=11pt,paper=a4]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{preamble}
\author{Mitja Daniel Krebs}
\date{\today}
\title{Congestion-Free Network Updates: Algorithms and Complexity}
\hypersetup{
 pdfauthor={Mitja Daniel Krebs},
 pdftitle={Congestion-Free Network Updates: Algorithms and Complexity},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 28.2 (Org mode 9.6.1)}, 
 pdflang={English}}
\begin{document}

@@

\begin{table}[htbp]
\caption{Table of notations}
\centering
\begin{tabular}{ll}
\hyperref[orge222158]{\(\leq_G\)} & The reachability relation for directed graph \(G\)\\[0pt]
\hyperref[orgc2eae0a]{\(b(v,P)\)} & The \(P\)-block containing vertex \(v\)\\[0pt]
\hyperref[orgef6d872]{\(P(b)\)} & The flow pair \(P\) such that block \(b\) is a \(P\)-block\\[0pt]
\hyperref[org5f5dd8d]{\(\mathcal{S}(b)\)} & The least vertex in block \(b\) w.r.t. reachability relation \(\leq_{P(b)^o\cup P(b)^u}\)\\[0pt]
\hyperref[org4de6595]{\(B^P(G)\)} & The set of \(P\)-blocks\\[0pt]
\hyperref[orgfb0a080]{\(B(G)\)} & The set of blocks\\[0pt]
\hyperref[org19b0df5]{\(\mathcal{B}(b)\)} & The round in which block \(b\) is updated\\[0pt]
\hyperref[org2c3faf3]{\(\mathcal{B}(v,P)\)} & The round in which block \(b(v,P)\) is updated\\[0pt]
\hyperref[orgd5c3eb8]{\(B_i\)} & The set of blocks updated before or in the \(i\)-th round\\[0pt]
\hyperref[orgf2fd7a2]{\(S(\pi)\)} & The set of which \(\pi\) is a permutation\\[0pt]
\hyperref[orgf2fd7a2]{\(\lvert\pi\rvert\)} & The number of elements of permutation \(\pi\)\\[0pt]
\hyperref[orgf2fd7a2]{\(\pi_i\)} & The \(i\)-th element of permutation \(\pi\)\\[0pt]
\hyperref[orgf2fd7a2]{\(\pi(x)\)} & The position of element \(x\) in permutation \(\pi\)\\[0pt]
\end{tabular}
\end{table}

\part{Preliminaries}
\label{sec:orgc92ee75}

\chapter{Blocks}
\label{sec:orgd8e1fcc}

\begin{notation}
For a directed graph \(G\), \(\leq_G=E(G)^*\) denotes the reachability relation of \(G\).
That is, for every two vertices \(u,v\in V(G)\), \(u\leq_Gv\) iff there is a path in \(G\) from \(u\) to \(v\).
\label{orge222158}
\end{notation}

@@

\begin{defn}
Let \(G=(V,E,\mathcal{P},s,t,c)\) be an update flow network and \(P\in\mathcal{P}\) be a flow pair.
Let \(v_1,\dots,v_{\ell}\) be the set \(V(P^o\cap P^u)\) ordered w.r.t. \(\leq_{P^o\cup P^u}\).
For every \(i\in[\ell-1]\), we define the \(i\)-th \emph{\(P\)-block} as \(b_i^P=\{v\mid v_i\leq_{P^o\cup P^u}v\leq_{P^o\cup P^u}v_{i+1}\}\).
\label{org39bfbe2}
\end{defn}

@@

\begin{notation}
For a flow pair \(P\) and a vertex \(v\in V(P)\), \(b(v,P)\) denotes the \(P\)-block containing \(v\).
\label{orgc2eae0a}
\end{notation}

\begin{notation}
For a block \(b\), \(P(b)\) denotes the flow pair \(P\) such that \(b\) is a \(P\)-block.
\label{org5f5dd8d}
\end{notation}

\begin{notation}
For a block \(b\), \(\mathcal{S}(b)\) denotes the \emph{start} of \(b\), that is, the least vertex in \(b\) w.r.t. \(\leq_{P(b)^o\cup P(b)^u}\).
\label{orgef6d872}
\end{notation}

\begin{notation}
For an update flow network \(G\) and a flow pair \(P\), \(B^P(G)\) denotes the set of \(P\)-blocks.
\label{org4de6595}
\end{notation}

\begin{notation}
For an update flow network \(G\), \(B(G)=\bigcup_{P\in\mathcal{P}}B^P(G)\) denotes the set of blocks.
\label{orgfb0a080}
\end{notation}

\chapter{Block Sequences}
\label{sec:org90a7988}

\begin{defn}
A \emph{block sequence} \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) is an ordered partition of the set of blocks.
\label{orgac79c0b}
\end{defn}

\begin{remark}
We may ignore all blocks containing less than three vertices.
\label{org22a2102}
\end{remark}

@@

\begin{notation}
For a block sequence \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) and a block \(b\), \(\mathcal{B}(b)\) denotes the index \(i\in[\ell]\) such that \(b\) is contained in \(\mathscr{B}_i\).
\label{org19b0df5}
\end{notation}

\begin{notation}
For a block sequence \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\), a flow pair \(P\), and a vertex \(v\in V(P)\), \(\mathcal{B}(v,P)=\mathcal{B}(b(v,P))\) denotes the index \(i\in[\ell]\) such that block \(b(v,P)\) is contained in \(\mathscr{B}_i\).
\label{org2c3faf3}
\end{notation}

\begin{notation}
For a block sequence \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) and an index \(i\in[\ell]\), \(B_i=\bigcup_{j\leq i}\mathscr{B}_j\) denotes the set of blocks updated before or in the \(i\)-th round.
\label{orgd5c3eb8}
\end{notation}

\begin{defn}
Let \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) be a block sequence.
For a flow pair \(P\), an edge \((u,v)\in E(P^o\cup P^u)\), and an index \(i\in[\ell]\), the \emph{activation label} \(\alpha_P((u,v),B_i)\) is defined as follows:
\[\alpha_P((u,v),B_i)=
\begin{cases}
\mathrm{active} & \text{if }(u,v)\in E(P^o)\text{ and }b(u,P)\notin B_i\\
\mathrm{active} & \text{if }(u,v)\in E(P^u)\text{ and }b(u,P)\in B_i\\
\mathrm{inactive} & \text{otherwise}.
\end{cases}\]
\label{org2ee7059}
\end{defn}

@@

\item If \((u,v)\in E(P^u\setminus P^o)\), then
\[\alpha_P((u,v),B_i)=
   \begin{cases}
   \mathrm{active} & i\geq\mathcal{B}(u,P)\\
   \mathrm{inactive} & i<\mathcal{B}(u,P).
   \end{cases}\]
\end{enumerate}
\label{org7e3e855}
\end{lem}

\begin{notation}
For a flow pair \(P\) and a \(P\)-block \(b\), \(U(b)=\{(v,P)\mid v\in b\}\) denotes the set of updates induced by \(b\).
Moreover, for a set \(B\) of blocks, \(U(B)=\bigcup_{b\in B}U(b)\) denotes the set of updates induced by \(B\).
\label{orgcfea092}
\end{notation}

@@

\begin{lem}
Let \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) be a block sequence, \(P\) be a flow pair, \(e\in E(P^o\cup P^u)\), and \(i\in[\ell]\).
Then \(\alpha_P(e,B_i)=\mathrm{active}\) iff \(e\in E(T_{P,U(B_i)})\).
\label{orga656ff1}
\end{lem}

\begin{defn}
A block sequence \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) is \emph{feasible} if for every edge \(e\) and every index \(i\in[\ell]\),
\begin{equation}
\label{eq:orgaadab49}
c(e)\geq\sum_{P\in\mathcal{P}:\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}}d_P,
\end{equation}
where we define \(\mathscr{B}_0\) to be the empty set.
\label{orgde355a9}
\end{defn}

\begin{remark}
Let \(G\) be an update flow network with unit demand, that is, \(d_P=1\) for every flow pair \(P\), and let \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) be a block sequence.
Then, for every edge \(e\) and every index \(i\in[\ell]\), capacity constraint \ref{eq:orgaadab49} simplifies to:
\begin{align*}
c(e)
&\geq\sum_{P\in\mathcal{P}:\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}}d_P\\
&=\sum_{P\in\mathcal{P}:\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}}1\\
&=\lvert\{P\in\mathcal{P}\mid\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}\}\rvert.
\end{align*}
\label{orgd3eb958}
\end{remark}

@@

\begin{enumerate}
\item \label{itm:lem-update-flow-network-feasible-if-1}
The old flow network is feasible if capacity constraint \ref{eq:orgaadab49} is satisfied for every edge and \(i=1\).

\item \label{itm:lem-update-flow-network-feasible-if-2}
The updated flow network is feasible if capacity constraint \ref{eq:orgaadab49} is satisfied for every edge and \(i=\ell\).
\end{enumerate}
\label{org51c1593}
\end{lem}

@@

\paragraph{\ref{itm:lem-update-flow-network-feasible-if-1}.}
Suppose capacity constraint \ref{eq:orgaadab49}  is satisfied for \(e\) and \(i=1\).
Then, since \(\mathscr{B}_0=\emptyset\), and by definitions of \(B_i\) and \(\alpha_P\):
\begin{align*}
c(e)
&\geq\sum_{P\in\mathcal{P}:\alpha_P(e,B_0)=\mathrm{active}\text{ or }\alpha_P(e,B_1)=\mathrm{active}}d_P\\
&\geq\sum_{P\in\mathcal{P}:\alpha_P(e,B_0)=\mathrm{active}}d_P\\
&=\sum_{P\in\mathcal{P}:e\in E(P^o)}d_P.
\end{align*}
\paragraph{\ref{itm:lem-update-flow-network-feasible-if-2}.}
Suppose capacity constraint \ref{eq:orgaadab49}  is satisfied for \(e\) and \(i=\ell\).
Then, since \(\mathcal{B}\) partitions the set of blocks, and by definitions of \(B_i\) and \(\alpha_P\):
\begin{align*}
c(e)
&\geq\sum_{P\in\mathcal{P}:\alpha_P(e,B_{\ell-1})=\mathrm{active}\text{ or }\alpha_P(e,B_{\ell})=\mathrm{active}}d_P\\
&\geq\sum_{P\in\mathcal{P}:\alpha_P(e,B_{\ell})=\mathrm{active}}d_P\\
&=\sum_{P\in\mathcal{P}:e\in E(P^u)}d_P.
\end{align*}
\end{proof}

\begin{corollary}
There is a feasible block sequence iff there is a feasible update sequence.
\label{org4ad2b7f}
\end{corollary}

\chapter{Block Permutations}
\label{sec:org962ccce}

\begin{notation}
For a set \(S\) and a permutation \(\pi=(x_1,\dots,x_{\ell})\) of \(S\),

\begin{enumerate}
\item \label{itm:notation-permutation-1}
\(S(\pi)\) denotes the set \(S\) of which \(\pi\) is a permutation;

\item \label{itm:notation-permutation-2}
\(\lvert\pi\rvert\) denotes the number \(\ell\) of elements of \(\pi\);

\item \label{itm:notation-permutation-3}
for an index \(i\in[\ell]\), \(\pi_i\) denotes the \(i\)-th element \(x_i\) of \(\pi\); and

\item \label{itm:notation-permutation-4}
for an element \(x\in S\), \(\pi(x)\) denotes the index \(i\) such that \(\pi_i=x\).
\end{enumerate}
\label{orgf2fd7a2}
\end{notation}

\begin{notation}
For two permutations \(\pi_1,\pi_2\), \(\mathrm{core}(\pi_1,\pi_2)=S(\pi_1)\cap S(\pi_2)\) denotes the \emph{core} of \(\pi_1\) and \(\pi_2\).
\label{org5d4037a}
\end{notation}

\begin{defn}
Let \(\pi\) be a permutation and \(\pi'\) be a subsequence of \(\pi\). Then:

\begin{enumerate}
\item \(\pi\) is an \emph{extension} of \(\pi'\) to \(S(\pi)\supseteq S(\pi')\); and

\item \(\pi'\) is the \emph{restriction} of \(\pi\) to \(S(\pi')\subseteq S(\pi)\).
\end{enumerate}
\label{orgc3d1fc9}
\end{defn}

\begin{itemize}
\item[{$\square$}] Should we define subsequence?
\end{itemize}

\begin{defn}
Two permutations \(\pi_1,\pi_2\) are \emph{consistent} if the restrictions of \(\pi_1\) and \(\pi_2\) to \(\mathrm{core}(\pi_1,\pi_2)\) are equal.
\label{org57f7c44}
\end{defn}

\begin{defn}
Let \(\pi_1,\pi_2\) be two consistent permutations.
A permutation \(\pi\) is a \emph{union} of \(\pi_1\) and \(\pi_2\) if \(\pi\) is an extension of both \(\pi_1\) and \(\pi_2\) to \(S(\pi_1)\cup S(\pi_2)\).
\label{org7053c59}
\end{defn}

\begin{defn}
Let \(X\subseteq E(G)\) be a set of edges.
A permutation \(\pi=(b_1,\dots,b_{\ell})\) of blocks is \emph{congestion free} w.r.t. \(X\) if for every edge \(e\in X\) and every index \(i\in[\ell]\),
\begin{equation}
\label{eq:org2d9cd7b}
c(e)\geq \sum_{P\in\mathcal{P}:\alpha_P(e,B_i)=\mathrm{active}}d_P,
\end{equation}
where \(B_i=\bigcup_{j\leq i}\{b_i\}\).
\label{org7c90f2d}
\end{defn}

\begin{itemize}
\item[{$\square$}] Remark why we may overload notation \(B_i\).
\end{itemize}

\begin{lem}
Let \(X\subseteq E(G)\) be a set of edges and \(\pi=(b_1,\dots,b_{\ell})\) be a permutation of blocks.
Then \(\pi\) is congestion free w.r.t. \(X\) iff the block sequence \(\mathcal{B}=(\{b_1\},\dots,\{b_{\ell}\})\) induced by \(\pi\) is feasible w.r.t. \(X\).
\label{org5cb530f}
\end{lem}

\begin{itemize}
\item[{$\square$}] The block sequence induced by \(\pi\) is not defined unless \(S(\pi)=B(G)\).

\item[{$\square$}] Feasible w.r.t. \(X\) is not defined unless \(X=E(G)\).
\end{itemize}

\begin{proof}
Let \(X\subseteq E(G)\) be a set of edges, \(\pi=(b_1,\dots,b_{\ell})\) be a permutation, and \(e\in X\) be an edge.

\paragraph{Only-if part.}
Let \(i\in[\ell]\).
If capacity constraint \ref{eq:orgaadab49} is satisfied for the block sequence \(\mathcal{B}=(\{b_1\},\dots,\{b_{\ell}\})\) induced by \(\pi\), \(e\), and \(i\), then capacity constraint \ref{eq:org2d9cd7b} is satisfied for \(\pi\), \(e\), and \(i\):
\[
\sum_{P\in\mathcal{P}:\alpha_P(e,B_i)=\mathrm{active}}d_P\leq\sum_{P\in\mathcal{P}:\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}}d_P\leq c(e)
\]

\paragraph{If part.}
We show the contrapositive.
Suppose capacity contraint \ref{eq:orgaadab49} is violated for some \(i\in[\ell]\).
We show that then capacity constraint \ref{eq:org2d9cd7b} is violated for \(i\) or \(i-1\).
Notice that if \(i=1\), then the latter contradicts the feasibility of update flow network \(G\).

Since block \(b_i\) is the only block updated in round \(i\), we have that for every block \(b\neq b_i\), \(b\in B_i\) iff \(b\in B_{i-1}\).
Hence for every flow pair \(P\in\mathcal{P}\setminus\{P(b_i)\}\), we have \(\alpha_P(e,B_i)=\mathrm{active}\) iff \(\alpha_P(e,B_{i-1})=\mathrm{active}\).
For flow pair \(P(b_i)\), we consider the cases \(\alpha_{P(b_i)}(e,B_i)=\mathrm{active}\) and \(\alpha_{P(b_i)}(e,B_{i-1})=\mathrm{active}\) separately.
(Note that if neither \(\alpha_{P(b_i)}(e,B_i)=\mathrm{active}\) nor \(\alpha_{P(b_i)}(e,B_{i-1})=\mathrm{active}\), then demand \(d_{P(b_i)}\) contributes to neither sum.)
If \(\alpha_{P(b_i)}(e,B_i)=\mathrm{active}\), then capacity constraint \ref{eq:org2d9cd7b} is violated for \(i\):
\begin{align*}
\sum_{P\in\mathcal{P}:\alpha_P(e,B_i)=\mathrm{active}}d_P&=\\
\sum_{P\in\mathcal{P}\setminus\{P(b_i)\}:\alpha_P(e,B_i)=\mathrm{active}}d_P+\sum_{P\in\{P(b_i)\}:\alpha_P(e,B_i)=\mathrm{active}}d_P&=\\
\sum_{P\in\mathcal{P}\setminus\{P(b_i)\}:\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}}d_P&\\
+\sum_{P\in\{P(b_i)\}:\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}}d_P&=\\
\sum_{P\in\mathcal{P}:\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}}d_P&>
c(e)
\end{align*}
If \(\alpha_{P(b_i)}(e,B_{i-1})=\mathrm{active}\), then capacity constraint \ref{eq:org2d9cd7b} is violated for \(i-1\):
\begin{align*}
\sum_{P\in\mathcal{P}:\alpha_P(e,B_{i-1})=\mathrm{active}}d_P&=\\
\sum_{P\in\mathcal{P}\setminus\{P(b_i)\}:\alpha_P(e,B_{i-1})=\mathrm{active}}d_P+\sum_{P\in\{P(b_i)\}:\alpha_P(e,B_{i-1})=\mathrm{active}}d_P&=\\
\sum_{P\in\mathcal{P}\setminus\{P(b_i)\}:\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}}d_P&\\
+\sum_{P\in\{P(b_i)\}:\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}}d_P&=\\
\sum_{P\in\mathcal{P}:\alpha_P(e,B_{i-1})=\mathrm{active}\text{ or }\alpha_P(e,B_i)=\mathrm{active}}d_P&>
c(e)
\end{align*}
\end{proof}

\begin{defn}
Let \(X\subseteq V(G)\) be a set of vertices.
A permutation \(\pi\) is an \(X\)-\emph{permutation} if:

\begin{enumerate}
\item \(S(\pi)=\bigcup_{P\in\mathcal{P},v\in X}\{b(v,P)\}\); and

\item \(\pi\) is congestion free w.r.t. \(E(X)\).
\end{enumerate}
\label{org72d1452}
\end{defn}

\part{\(\textbf{NP}\)-Hardness for \(k=3\)}
\label{sec:org53d400c}

@@

\begin{thm}
The \(k\)-network flow update problem is \(\textbf{NP}\)-hard for \(k=3\).
\label{org643d469}
\end{thm}

@@

\begin{thm}
The \(k\)-network flow update problem, where every edge is used by at most three flow pairs, is \(\textbf{NP}\)-hard for \(k=10\).
\label{org3e793cd}
\end{thm}

Then, we will (repeatedly) apply the following lemma to the flow update network we will have constructed in the proof of Theorem \ref{org3e793cd} to reduce the number of flow pairs from \(10\) to \(3\).

@@

Then there is an update flow network \(\tilde{G}\) with \(k-1\) flow pairs such that (\(\lvert\tilde{G}\rvert=O(\lvert G\rvert)\) and) there is a feasible block sequence \(\mathcal{B}=(\mathscr{B}_1,\dots,\mathscr{B}_{\ell})\) for \(G\) iff there is a feasible block sequence \(\tilde{\mathcal{B}}=(\tilde{\mathscr{B}}_1,\dots,\tilde{\mathscr{B}}_{\ell})\) for \(\tilde{G}\).
\label{org1685041}
\end{lem}

@@

\chapter{\(\textbf{NP}\)-Hardness for the Special Case}
\label{sec:orgd040ef0}

The proof of Theorem \ref{org3e793cd} is via reduction from \(\textsf{4-SAT}\) and is based on the \(\textbf{NP}\)-hardness proof for \(k=6\) in (Amiri, Saeed A. and Dudycz, Szymon and Parham, Mahmoud and Schmid, Stefan and Wiederrecht, Sebastian, 2019).

@@

\section{The Reduction}
\label{sec:orgab5d1b8}

@@

We now proceed with the detailed specification of clause gadget \(C^i\) (see Figure \ref{fig:orgb090ce4}).
